---
title: "OSM Data Description"
author: "Grégoire Lurton"
date: "March 18, 2016"
output: html_document
---

First level of analysis, description of Open Street Map data we extracted for Nigeria

```{r setup, include=FALSE}
library(maptools)
library(knitr)
library(pander)
library(plyr)

NigeriaShp <-  readShapePoly('J://Project/phc/nga/dhis/LGAMap.shp')
osm_data <-  readShapePoints('data/OSMDataNigeria.shp')
```

## Mapped health facilities

Two sources have reported health facilities :

1. the ehealthafrica project has mapped facilities in Kano , Bauchi and Borno

2. the NMIS / PEPFAR project has mapped facilities in Edo Sate

We first look at the distribution of these facilties :
 
```{r get and clean}
health_projects = osm_data[grepl(x = tolower(osm_data$source) , pattern = 'ehealthafrica|nmis|pepfar') , ]


## Getting ehealthAfrica data
ehealth_facilities <- health_projects[!is.na(health_projects$health_fac) , ]
ehealth_facilities$addr_state <- as.character(ehealth_facilities$addr_state)
ehealth_facilities$addr_state[is.na(ehealth_facilities$addr_state)] <- 'Borno'
ehealth_facilities$addr_state[ehealth_facilities$addr_state == 'KN'] <- 'Kano'

duplicates_ehealth <- ddply(ehealth_facilities@data , .(name , addr_state) , nrow)
duplicates_ehealth <- subset(duplicates_ehealth , V1 > 1)
ehealth_facilities <- ehealth_facilities[!(ehealth_facilities$name %in% duplicates_ehealth$name) ,]


## Getting NMIS / PEPFAR data
nmis_facilities <- health_projects[health_projects$source %in% c('NMIS','PEPFAR') , ]
## Dropping outliers
nmis_facilities <- nmis_facilities[as.numeric(as.character(nmis_facilities$lat)) < min(as.numeric(as.character(ehealth_facilities$lat))) ,]

duplicates_nmis <- ddply(nmis_facilities@data , .(primary_na) , nrow)
duplicates_nmis <- subset(duplicates_nmis , V1 > 1)
nmis_facilities <- nmis_facilities[!(nmis_facilities$primary_na %in% duplicates_nmis$primary_na) ,]
```


```{r map facilties osm}
plot(NigeriaShp)
plot(ehealth_facilities , col = factor(ehealth_facilities$addr_state) , add = TRUE)
plot(nmis_facilities , col = 'blue' , add = TRUE)
```

The distribution of facilities mapped by ehealthafrica is as follow :

```{r table ehealthAfrica}
pander(table(ehealth_facilities$addr_state , as.character(ehealth_facilities$health_fac)))
pander(table(as.character(nmis_facilities$source)))
```

